Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-20 arxiv/proof_main.tex
revise

\medskip
\noindent{\bf Concentration estimates.} Now for the right-hand side of equation \eqref{0self_Gii}, notice that the resolvent minor $G^{(i)}$ is defined such that it is independent of the entries $Z^{(1)}_{\mu i}$ and $Z^{(2)}_{\nu i}$. Hence by the concentration inequalities in Lemma \ref{largedeviation}, we have that the  right-hand side of equation \eqref{0self_Gii} concentrates around the partial expectation over the entries $\{Z^{(1)}_{\mu i}: \mu \in \cal I_1 \}\cup \{Z^{(2)}_{\nu i}: \nu \in \cal I_2\}$, i.e., with high probability,
\begin{align*}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu \in \mathcal I_1}  G^{\left( i \right)}_{\mu\mu} - \frac{1}{n} \sum_{\mu\in \mathcal I_2} G^{\left( i \right)}_{\mu\mu} +\oo(1)= - z - \lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1^{(i)}(z)-  \frac{\rho_2}{\rho_1+\rho_2} m_2^{(i)}(z)+\oo(1),
\end{align*}
where we used the definition of $m_1^{(i)}$ and $m_2^{(i)}$ in equation \eqref{defm} with $G$ replaced by $G^{(i)}$. Intuitively, since we have removed only one column and one row out of the $(p+n)$ columns and rows in $H$, $m_1^{(i)}$ and $m_2^{(i)}$ should be close to the original $m_1$ and $m_2$. Hence we obtain from the above equation that
\begin{align}\label{1self_Gii}
 G_{ii}  = -\left( z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1 +  \frac{\rho_2}{\rho_1+\rho_2}m_2+\oo(1)\right)^{-1}.
\end{align}
Similarly, we can obtain from equation \eqref{0self_Gmu1} that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{1self_Gmu} G_{\mu \mu }=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},\ee
with high probability. The rigorous derivation of the above concentration estimates will be given in the proof of Lemma \ref{lemm_selfcons_weak}; see equations \eqref{self_Gii}-\eqref{erri}.

\medskip
\noindent\textbf{Deriving the self-consistent equations.} Now taking average of equation \eqref{1self_Gmu}, we obtain that
\be\label{2self_Gmu} m_1= \frac{1}{n_1}\sum_{\mu \in \cal I_1}G_{\mu\mu}=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad m_2=\frac{1}{n_2}\sum_{\nu \in \cal I_2}G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},
\ee
with high probability. As a byproduct, comparing equations \eqref{1self_Gmu} and \eqref{2self_Gmu}, we obtain that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{2.5self_Gmu} G_{\mu \mu }=m_1 +\oo(1),\quad G_{\nu\nu}=m_2+\oo(1),\ee
with high probability. Together with the definition of $m$ and $m_0$ in equation \eqref{defm}, the two equations in equation \eqref{2self_Gmu} give that
\be\label{3self_Gmu}  \frac1{m_1}= -1- \frac{1}{n} \sum_{i=1}^p \lambda_i^2 G_{ii}+ \oo(1),\quad \frac1{m_2}=-1-\frac{1}{n} \sum_{i=1}^p G_{ii}  + \oo(1),\ee
with high probability. Plugging equation \eqref{1self_Gii} into equation \eqref{3self_Gmu}, we obtain that
\be\label{approximate m1m2}
\begin{split}
& \frac1{m_1}= -1+ \frac{1}{n} \sum_{i=1}^p \frac{\lambda_i^2 }{ z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)}+ \oo(1),\\
& \frac1{m_2}=-1+\frac{1}{n} \sum_{i=1}^p \frac{1}{ z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)}  + \oo(1),
\end{split}
\ee
with high probability, which give a system of approximate self-consistent equations for $(m_1,m_2)$. 
A rigorous derivation of these two equations will be given in the proof of Lemma \ref{lemm_selfcons_weak}. Compare \eqref{approximate m1m2} to the deterministic self-consistent equations in equation \eqref{selfomega_a}, one can observe that we should have
\be\label{approx m12 add}
(m_1,m_2) =\left(-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z),-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\right)+\oo(1) \quad \text{with high probability. }
\ee
To justify this identity rigorously, we need to know that the self-consistent equations are stable, that is, a small perturbation of the equations leads to a small perturbation of the solution. This will be given by Lemma \ref{lem_stabw}.

\medskip
\noindent\textbf{Deriving the matrix limit.}  Inserting the approximate identity \eqref{approx m12 add} into equations \eqref{1self_Gii} and \eqref{2.5self_Gmu}, we get that for  $i \in \cal I_0$, $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
$$G_{ii}(z)=(-z +\lambda_i^2 a_{1}(z) + a_2(z)+\oo(1)^{-1},\quad G_{\mu\mu}=-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z)+\oo(1),\quad G_{\nu\nu}=-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)+\oo(1),$$
with high probability. These explain the diagonal entries of $\Gi$ in equation \eqref{defn_piw}. For the off-diagonal entries, they are close to zero due to concentration. For example, for $i\ne j\in \cal I_1$, by Schur complement formula in equation (\ref{resolvent3}), we have
$$G_{ij}=-G_{ii}\Big({\lambda_i}{n^{-1/2}}\sum_{\mu \in \cal I_1} Z^{(1)}_{\mu i} G^{(i)}_{\mu j} + {n^{-1/2}}\sum_{\mu \in \cal I_2} Z^{(2)}_{\mu i} G^{(i)}_{\mu j} \Big).$$
Using Lemma \ref{largedeviation}, we can show that $n^{-1/2}\sum_{\mu \in \cal I_1} Z^{(1)}_{\mu i} G^{(i)}_{\mu j}$ and $n^{-1/2}\sum_{\mu \in \cal I_2} Z^{(2)}_{\mu i} G^{(i)}_{\mu j}$ are both close to zero. The other off-diagonal entries can be bounded in the same way. The bound on the off-diagonal entries will be given by Lemma \ref{Z_lemma}.

>>>>

\medskip
\noindent{\bf Concentration estimates.} Now for the right-hand side of equation \eqref{0self_Gii}, notice that the resolvent minor $G^{(i)}$ is defined such that it is independent of the entries $Z^{(1)}_{\mu i}$ and $Z^{(2)}_{\nu i}$. Hence by the concentration inequalities in Lemma \ref{largedeviation}, we have that the  right-hand side of equation \eqref{0self_Gii} concentrates around the partial expectation over the entries $\{Z^{(1)}_{\mu i}: \mu \in \cal I_1 \}\cup \{Z^{(2)}_{\nu i}: \nu \in \cal I_2\}$, i.e., with overwhelming probability,
\begin{align*}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu \in \mathcal I_1}  G^{\left( i \right)}_{\mu\mu} - \frac{1}{n} \sum_{\mu\in \mathcal I_2} G^{\left( i \right)}_{\mu\mu} +\oo(1)= - z - \lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1^{(i)}(z)-  \frac{\rho_2}{\rho_1+\rho_2} m_2^{(i)}(z)+\oo(1),
\end{align*}
where we used the definition of $m_1^{(i)}$ and $m_2^{(i)}$ in equation \eqref{defm} with $G$ replaced by $G^{(i)}$. Intuitively, since we have removed only one column and one row out of the $(p+n)$ columns and rows in $H$, $m_1^{(i)}$ and $m_2^{(i)}$ should be close to the original $m_1$ and $m_2$. Hence we obtain from the above equation that
\begin{align}\label{1self_Gii}
 G_{ii}  = -\left( z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1 +  \frac{\rho_2}{\rho_1+\rho_2}m_2+\oo(1)\right)^{-1}.
\end{align}
Similarly, we can obtain from equation \eqref{0self_Gmu1} that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{1self_Gmu} G_{\mu \mu }=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},\ee
with overwhelming probability. The rigorous derivation of the above concentration estimates will be given in the proof of Lemma \ref{lemm_selfcons_weak}; see equations \eqref{self_Gii}-\eqref{erri}.

\medskip
\noindent\textbf{Deriving the self-consistent equations.} Now taking average of equation \eqref{1self_Gmu}, we obtain that
\be\label{2self_Gmu} m_1= \frac{1}{n_1}\sum_{\mu \in \cal I_1}G_{\mu\mu}=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad m_2=\frac{1}{n_2}\sum_{\nu \in \cal I_2}G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},
\ee
with overwhelming probability. As a byproduct, comparing equations \eqref{1self_Gmu} and \eqref{2self_Gmu}, we obtain that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{2.5self_Gmu} G_{\mu \mu }=m_1 +\oo(1),\quad G_{\nu\nu}=m_2+\oo(1),\ee
with overwhelming probability. Together with the definition of $m$ and $m_0$ in equation \eqref{defm}, the two equations in equation \eqref{2self_Gmu} give that
\be\label{3self_Gmu}  \frac1{m_1}= -1- \frac{1}{n} \sum_{i=1}^p \lambda_i^2 G_{ii}+ \oo(1),\quad \frac1{m_2}=-1-\frac{1}{n} \sum_{i=1}^p G_{ii}  + \oo(1),\ee
with overwhelming probability. Plugging equation \eqref{1self_Gii} into equation \eqref{3self_Gmu}, we obtain that
\be\label{approximate m1m2}
\begin{split}
& \frac1{m_1}= -1+ \frac{1}{n} \sum_{i=1}^p \frac{\lambda_i^2 }{ z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)}+ \oo(1),\\
& \frac1{m_2}=-1+\frac{1}{n} \sum_{i=1}^p \frac{1}{ z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)}  + \oo(1),
\end{split}
\ee
with overwhelming probability, which give a system of approximate self-consistent equations for $(m_1,m_2)$. 
A rigorous derivation of these two equations will be given in the proof of Lemma \ref{lemm_selfcons_weak}. Compare \eqref{approximate m1m2} to the deterministic self-consistent equations in equation \eqref{selfomega_a}, one can observe that we should have
\be\label{approx m12 add}
(m_1,m_2) =\left(-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z),-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\right)+\oo(1) \quad \text{with overwhelming probability. }
\ee
To justify this identity rigorously, we need to know that the self-consistent equations are stable, that is, a small perturbation of the equations leads to a small perturbation of the solution. This will be given by Lemma \ref{lem_stabw}.

\medskip
\noindent\textbf{Deriving the matrix limit.}  Inserting the approximate identity \eqref{approx m12 add} into equations \eqref{1self_Gii} and \eqref{2.5self_Gmu}, we get that for  $i \in \cal I_0$, $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
$$G_{ii}(z)=(-z +\lambda_i^2 a_{1}(z) + a_2(z)+\oo(1)^{-1},\quad G_{\mu\mu}=-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z)+\oo(1),\quad G_{\nu\nu}=-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)+\oo(1),$$
with overwhelming probability. These explain the diagonal entries of $\Gi$ in equation \eqref{defn_piw}. For the off-diagonal entries, they are close to zero due to concentration. For example, for $i\ne j\in \cal I_1$, by Schur complement formula in equation (\ref{resolvent3}), we have
$$G_{ij}=-G_{ii}\Big({\lambda_i}{n^{-1/2}}\sum_{\mu \in \cal I_1} Z^{(1)}_{\mu i} G^{(i)}_{\mu j} + {n^{-1/2}}\sum_{\mu \in \cal I_2} Z^{(2)}_{\mu i} G^{(i)}_{\mu j} \Big).$$
Using Lemma \ref{largedeviation}, we can show that $n^{-1/2}\sum_{\mu \in \cal I_1} Z^{(1)}_{\mu i} G^{(i)}_{\mu j}$ and $n^{-1/2}\sum_{\mu \in \cal I_2} Z^{(2)}_{\mu i} G^{(i)}_{\mu j}$ are both close to zero. The other off-diagonal entries can be bounded in the same way. The bound on the off-diagonal entries will be proved rigorously in Lemma \ref{Z_lemma}.

### 2020-09-20 arxiv/setup.tex
wip
revise

\subsection{Notations}
We shall say an event $\Xi$ holds with high probability (w.h.p.) if for any fixed $D>0$, $\P(\Xi)\ge 1- p^{-D}$ for large enough $p$. Moreover, we shall use $\oo(1)$ to mean a small positive quantity that converges to 0 as $p\to \infty$.
For a matrix $X$, let $\lambda_{\min}(X)$ denote its smallest singular value and $\norm{X}$ denote its spectral norm.
We shall refer to random matrices of the form $\frac {X^\top X} n$ as sample covariance matrices following the standard notations in high-dimensional statistics.
%\subsection{Bias and Variance}

@@

\begin{fact}\label{lem_minv}
	Let $c$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
	%Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	%Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	With high probability over the randomness of $X$, we have that
	\begin{enumerate}
		\item The sample covariance matrix $\frac{X^{\top}X}{n}$ is full rank.
		\item $\bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{p}{n - p} + \OO(p^{-c})$, cf. Theorem 2.4 in \citet{isotropic}.
	\end{enumerate}
\end{fact}

>>>>

\subsection{Notations}
We shall use $\oo(1)$ to mean a small positive quantity that converges to 0 as $p\to \infty$. 
We shall say an event $\Xi$ holds with high probability (w.h.p.) if $\P(\Xi)\ge 1- \oo(1)$, while we say $\Xi$ holds with overwhelming probability (w.o.p.) if for any large constant $D>0$, $\P(\Xi)\ge 1- p^{-D}$ for large enough $p$. For a matrix $X$, let $\lambda_{\min}(X)$ denote its smallest singular value and $\norm{X}$ denote its spectral norm.
We shall refer to random matrices of the form $\frac {X^\top X} n$ as sample covariance matrices following the standard notations in high-dimensional statistics.
%\subsection{Bias and Variance}

@@

\begin{fact}\label{lem_minv}
	Let $c$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
	%Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	%Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	With high probability over the randomness of $X$, we have that
	\begin{enumerate}
		\item The sample covariance matrix $\frac{X^{\top}X}{n}$ is full rank.
		\item $\bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{p}{n - p} + \OO(p^{-c})$, cf. Theorem 2.4 in \citet{isotropic}.
		\item The singular values of $Z^{\top}Z$ are greater than $(\sqrt{n} - \sqrt{p})^2 - n \cdot p^{-c}$ and less than $(\sqrt{n} + \sqrt{p})^2 + n \cdot p^{-c}$.
	\end{enumerate}
\end{fact}
